---
title: "DBI"
---

## Export to CSV and import to R

```{r, eval=FALSE}
res <- read.csv("/tmp/3000_hit_club.csv")

library(ggplot2)
ggplot(data = res, aes(x = reorder(playerID, tH), y = tH)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Clunky, error-prone, not automated, not reproducible

## Connect directly using DBI

(https://github.com/rstats-db/DBI)

Connects seamlessly to SQLite, MySQL, PostgreSQL, MonetDB, etc.

```{r}
library(DBI)
conn <- dbConnect(RMySQL::MySQL(), dbname = "lahman", user = "r-user", password = "mypass")
```

Retrieve results using `dbGetQuery()`.

```{r}
sql <- "
SELECT playerID, sum(H) as tH 
  FROM Batting
  GROUP BY playerID
  HAVING tH >= 3000
  ORDER BY tH desc;"
res <- dbGetQuery(conn, sql)
res
```

### Get Info

About a connection

```{r}
dbGetInfo(conn)
dbListFields(conn, name = "Batting")
```

About a database/schema

```{r}
dbListTables(conn)
```

About a table

```{r}
dbListFields(conn, name = "Batting")
```

### Read tables

```{r}
Teams <- dbReadTable(conn, "Teams")
str(Teams)
```

### Many other things...

Many other SQL operations including writing tables, performing transactions, etc. are possible through `DBI`.

