---
title: "dplyr and SQL"
---

```{r, echo=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(error = TRUE, message = FALSE, eval=FALSE)
```


## `dplyr` and SQL

```{r, message=FALSE}
library(dplyr)
library(DBI)
db <- src_mysql(user = "r-user", password = "mypass", dbname = "lahman")
dbListTables(db$con)
class(db)
str(db)
class(db$con)
```

## Question 1

```
SELECT playerID, sum(H) as tH 
  FROM Batting
  GROUP BY playerID
  HAVING tH >= 3000
  ORDER BY tH desc;
```

```{r}
batting <- tbl(db, "Batting")
res <- batting %>%
  group_by(playerID) %>%
  summarize(tH = sum(H)) %>%
  filter(tH >= 3000) %>%
  arrange(desc(tH))
```

```{r}
library(ggplot2)
ggplot(data = collect(res), aes(x = reorder(playerID, tH), y = tH)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## `dplyr` data verbs

[Introduction to `dplyr`](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html)

Based on SQL syntax:

* `select()` -> `SELECT`
* `mutate()` -> user-defined columns
* `summarize()` -> aggregated columns
* `left_join()` -> `LEFT JOIN`
* `filter()` -> `WHERE`
* `group_by()` -> `GROUP BY`
* `filter()` -> `HAVING`
* `arrange()` -> `ORDER BY`

## `dplyr` translates R code into SQL code

```{r}
res %>% 
  show_query()
```

## Pass SQL commands through R

```
SELECT b.playerID, CONCAT(m.nameLast, ', ', m.nameFirst) as Name, 
    CONCAT(min(yearID), '-', max(yearID)) as span, sum(H) as tH
  FROM Batting b
  LEFT JOIN Master m ON b.playerID = m.playerID 
  WHERE yearID BETWEEN 1980 AND 1989
  GROUP BY playerID
  ORDER BY tH desc
  LIMIT 0, 10;
```

This looks like valid R code:

```{r}
master <- tbl(db, "Master")
batting %>%
  left_join(master, by = "playerID") %>%
  filter(yearID >= 1980, yearID <= 1989) %>%
  group_by(playerID) %>%
  summarize(name = first(paste0(nameLast, ', ', nameFirst)), tH = sum(H)) %>%
  arrange(desc(tH))
```

But it doesn't work because MySQL doesn't have a command called `PASTE0`. Note that `dplyr` doen't know how to translate `paste0()`.

```{r, eval=TRUE}
translate_sql(paste0("hi"))
```

There are two workarounds for this: either write SQL code into the pipeline, or use `collect()` to bring the data back into R first. The former will be more efficient, but might look weird. 

```{r}
batting %>%
  left_join(master, by = "playerID") %>%
  filter(yearID >= 1980, yearID <= 1989) %>%
  group_by(playerID) %>%
  summarize(name = MAX(CONCAT(nameLast, ', ', nameFirst)), tH = sum(H)) %>%
  arrange(desc(tH))
```

The approach with `collect()` will be less efficient.

```{r}
master_lcl <- collect(master)
batting_lcl <- collect(batting)
batting_lcl %>%
  left_join(master_lcl, by = "playerID") %>%
  filter(yearID >= 1980, yearID <= 1989) %>%
  group_by(playerID) %>%
  summarize(name = max(paste0(nameLast, ', ', nameFirst)), tH = sum(H)) %>%
  arrange(desc(tH))
```

Notice the differences between `master_lcl` and `master`:

```{r}
class(master)
class(master_lcl)
print(object.size(master), units = "Mb")
print(object.size(master_lcl), units = "Mb")
```

## Plotting

```
SELECT yearID, lgID, sum(SO) / sum(AB + BB + HBP + SF + SH) as SO_rate
  FROM Batting 
  WHERE yearID >= 1954
  GROUP BY yearID, lgID;
```


```{r}
so_rates <- batting %>%
  filter(yearID >= 1954) %>%
  group_by(yearID, lgID) %>%
  summarize(so_rate = sum(SO) / sum(AB + BB + HBP + SF + SH)) %>%
  collect()
library(ggplot2)
ggplot(data = so_rates, aes(x = yearID, y = so_rate, color = lgID)) + 
  geom_line()
```

## Fitting models

```
SELECT teamID, yearID, W / (W + L) as WPct, R as RS, RA
  FROM Teams
  WHERE yearID >= 1954;
```

```{r}
pythag <- tbl(db, "Teams") %>%
  filter(yearID >= 1954) %>%
  mutate(wpct = W / (W + L), run_ratio = R / RA) %>%
  select(teamID, yearID, wpct, run_ratio) %>%
  collect()
ggplot(data = pythag, aes(x = run_ratio, y = wpct)) + 
  geom_point(alpha = 0.5)
library(mosaic)
fitModel(formula = wpct ~ 1 / (1 + (1 / run_ratio)^k), data = pythag) %>%
  coef()
```



