---
title: "Practice: Joining Tables"
---

```{r, echo=FALSE, message=FALSE} 
require(knitr)
opts_chunk$set(echo = FALSE, eval=FALSE)
```

```{r connect}
require(RPostgreSQL)
con <- dbConnect(PostgreSQL(), host="localhost", user="postgres", password="postgres", dbname="airlines")
```

#### Questions

#. Find the names of all of the carriers that left Bradley International Airport on June 15th, 2013. 

```{r}
dbGetQuery(con, statement = "SELECT c.* FROM flights f JOIN carriers c ON f.carrier = c.carrier WHERE month = 6 and day = 15 and origin = 'BDL';")
```

Use the `airlines` database to answer the following questions. 

#. Find all the carriers that fly out of Bradley International Airport and list them by number of flights. 

```{r}
dbGetQuery(con, statement = "SELECT carrier, sum(1) as N FROM flights WHERE origin = 'BDL' GROUP BY carrier ORDER BY N desc;")
```


#. Find all the **names** of all of the carriers that fly out of Bradley International Airport and list them by number of flights. 

```{r}
dbGetQuery(con, statement = "SELECT f.carrier, max(c.name) as Airline, sum(1) as N FROM flights f LEFT JOIN carriers c ON f.carrier = c.carrier WHERE origin = 'BDL' GROUP BY f.carrier ORDER BY N desc;")
```

#. What model of plane is most common at Bradley?

```{r}
dbGetQuery(con, statement = "SELECT manufacturer, model, sum(1) as N FROM flights f LEFT JOIN planes p ON f.tailnum = p.tailnum WHERE origin = 'BDL' GROUP BY manufacturer, model ORDER BY N desc;")
```

#. What model of plane is most common at Bradley? This time group the model numbers before the `-`.

```{r}
dbGetQuery(con, statement = "SELECT manufacturer, split_part(model, '-', 1) as myModel, sum(1) as N FROM flights f LEFT JOIN planes p ON f.tailnum = p.tailnum WHERE origin = 'BDL' GROUP BY manufacturer, myModel ORDER BY N desc;")
```
