---
title: "MySQL Workbench"
---

## Introduction to Lahman database

```{r, echo=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, eval = FALSE)
library(dplyr)
library(DBI)
db <- src_mysql(user = "r-user", password = "mypass", dbname = "lahman")
dbListTables(db$con)
```

```{r}
opts_chunk$set(engine.opts = "-t lahman")
```



## Motivating questions

#. Find all batters who have accumulated at least 3000 hits in their career.

```{mysql}
SELECT playerID, sum(H) as tH 
  FROM Batting
  GROUP BY playerID
  HAVING tH >= 3000
  ORDER BY tH desc;
```

#. Find all batters who have accumulated at least 500 home runs in their career.

```{mysql}
SELECT playerID, sum(HR) as tHR 
  FROM Batting
  GROUP BY playerID
  HAVING tHR >= 500
  ORDER BY tHR desc;
```

#. Find all batters who have accumulated at least 300 home runs and at least 300 stolen bases in their career. 

```{mysql}
SELECT playerID, sum(HR) as tHR, sum(SB) as tSB 
  FROM Batting
  GROUP BY playerID
  HAVING tHR >= 300 AND tSB >= 300
  ORDER BY tHR desc;
```

#. Find all pitchers who have accumulated at least 300 wins and at least 3000 strikeouts in their career. 

```{mysql}
SELECT playerID, sum(W) as tW, sum(SO) as tSO 
  FROM Pitching
  GROUP BY playerID
  HAVING tW >= 300 AND tSO >= 3000
  ORDER BY tW desc;
```

#. Find all batters with at least 50 home runs in a single season.


```{mysql}
SELECT playerID, yearID, sum(HR) as tHR
  FROM Batting
  GROUP BY playerID, yearID
  HAVING tHR >= 50
  ORDER BY tHR desc;
```


#. Find all batters who had at least 40 home runs and at least 40 stolen bases in a single season.


```{mysql}
SELECT playerID, yearID, sum(HR) as tHR, sum(SB) as tSB 
  FROM Batting
  GROUP BY playerID, yearID
  HAVING tHR >= 40 AND tSB >= 40
  ORDER BY tHR desc;
```

#. Find the names of the top 10 batters who had the most hits during the decade of the 1980s.

```{mysql}
SELECT b.playerID, CONCAT(m.nameLast, ', ', m.nameFirst) as Name, 
    CONCAT(min(yearID), '-', max(yearID)) as span, sum(H) as tH
  FROM Batting b
  LEFT JOIN Master m ON b.playerID = m.playerID 
  WHERE yearID BETWEEN 1980 AND 1989
  GROUP BY playerID
  ORDER BY tH desc
  LIMIT 0, 10;
```


#. Find the names of all batters who have accumulated at least 3000 hits **or** at least 500 home runs but are **not** in the Hall of Fame.

```{mysql}
SELECT b.playerID, CONCAT(m.nameLast, ', ', m.nameFirst) as Name, 
    CONCAT(min(b.yearID), '-', max(b.yearID)) as span, 
    sum(H) as tH, sum(HR) as tHR, max(h.inducted) as inducted
  FROM Batting b
  LEFT JOIN Master m ON b.playerID = m.playerID 
  LEFT JOIN (
    SELECT playerID, inducted 
    FROM HallOfFame 
    WHERE inducted = 'Y' AND votedBy = 'BBWAA'
  ) h ON h.playerId = b.playerID
  GROUP BY playerID
  HAVING tH >= 3000 OR tHR > 500
  ORDER BY tH desc;
```



#. Make a plot showing the average strikeout rate in the AL and NL, going back to 1954.

```{mysql}
SELECT yearID, lgID, sum(SO) / sum(AB + BB + HBP + SF + SH) as SO_rate
  FROM Batting 
  WHERE yearID >= 1954
  GROUP BY yearID, lgID;
```



#. Consider Bill James' famous ["Pythagorean" model](https://en.wikipedia.org/wiki/Pythagorean_expectation) for expected winning percentage. 
$$
  \hat{WPct} = \frac{1}{1 + (RA/RS)^k}, 
$$
where $WPct$ is team winning percentage and $RS$ and $RA$ are team runs scored and allowed, respectively. Find the optimal value of $k$ using data since 1954. 



```{mysql}
SELECT teamID, yearID, W / (W + L) as WPct, R as RS, RA
  FROM Teams
  WHERE yearID >= 1954;
```
