---
title: 'Resources'
---


* Using SQL:
    + Download and install [MySQL](http://www.mysql.com/download)
    + Download and install the [MySQL Workbench](http://dev.mysql.com/downloads/workbench/). This is the officially-supported GUI for working with MyQL. It is free, open-source, and cross-platform. 
    + `dplyr` provides a syntax for pulling MySQL data directly into R, but relies upon the [`RMySQL` package](http://cran.r-project.org/web/packages/RMySQL/index.html) to do so. This package is cross-platform, but is not always available as a binary. 
    + Installing `RMySQL`:
        ```{r, eval=FALSE}
        install.packages("RMySQL")
        ```
        If that doesn't work, try:
        
        ```{r, eval=FALSE}
        install.packages("RMySQL", type="source")
        ```
     + Read the [MySQL Documentation](https://dev.mysql.com/doc/refman/5.7/en/)!
* Books
    + [SQL in a Nutshell](http://www.amazon.com/SQL-Nutshell-OReilly-Kevin-Kline/dp/0596518846/ref=sr_1_1?s=books&ie=UTF8&qid=1433865473&sr=1-1&keywords=sql+in+a+nutshell)
